Supporting Information for Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-09-18 paper.tex
edits

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{taxonomy.png}
	\caption{
		Some spectroscopic methods for investigating (top) electronic-vibrational coupling and (bottom) e-v-x (x: e, v, e-v, e-v-e) coupling through different order nonlinear processes.
		The light-matter interactions are depicted using Wave Mixing Energy Level (WMEL) diagrams.\cite{RN286}
		Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. 
		The red box outlines the e-v coupling techniques (HDFG, HSFG) relevant to this manuscript.
	} %TODO: Fix vsfg
	\label{fig:comparisonwmel}
\end{figure*}

@@

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{drsive_spectrum.png}
	\caption{Contributions of $A, B$ (\autoref{ABterms_DR}) to HDFG spectrum for a simple two harmonic well system.
		(a) Potential energy surfaces for a two-well system, (b) 1D HDFG spectrum for $\Delta = 0.5$, $\omega_1 = \omega_{g1, g0}$, (c,d) real and imaginary parts of terms plotted in (b), respectively.
		All spectra are referenced to $\max{(A+B)}$. 
		It is assumed that $\omega_2 = \omega_3$.
		Note that (b) plots magnitudes of the labeled quantities.
		The vibrational states on the $|g)$ and $|e)$ manifolds are spaced 1600 cm$^{-1}$ apart, with the vibronic states $\ket{e,\tilde{v}'}$ given linewidths of 400 cm$^{-1}$ and $\hbar \omega_{eg}$ = 30000 cm$^{-1}$.
		Dotted lines in (b) denote the $\tilde{v}'$ = 0, 1, 2 vibronic resonances. 
		The vibronic one and two photon absorption operators are scaled such that $\abs{B}/\abs{A} \sim$ 0.1, following Chung and Ziegler. \cite{Ziegler1988}
		The Franck-Condon and Herzberg-Teller contributions to each of these terms are found in \autoref{T:contributions}.}
	\label{fig:doubres_spec}
\end{figure*}

@@

In the post-resonance region, the amplitude takes on an unexpected lineshape.
This arises due to interference between the real and imaginary parts of $\gamma$ (\autoref{fig:doubres_spec}c, d).
The abnormal lineshapes in both terms arises due to the sign of the Franck-Condon and Herzberg-Teller integrals that drive the output of the vibronic transitions for $\Delta = 0.5$ (\autoref{fcht}).
Such effects have been seen elsewhere in spontaneous, surface enhanced hyper-Raman studies. \cite{Milojevich2011}
Since $\Delta$ can take on a variety of values given a specific system, in the supporting information we demonstrate the impact of $\Delta$ on the HDFG response using previously reported values for Raman active vibrations in isoprene. \cite{Myers1982}
Clearly, specific $\Delta$ values have a significant impact on the

@@

\section{Conclusions}\label{conclusion}
\textbf{NEED TO READ THIS SECTION!}

Coherent vibrational, hyper-Raman coherent four wave mixing spectroscopies are identified and discussed.
Hyper difference frequency generation (HDFG) possesses infrared absorption and vibronic hyper-Raman scattering selection rules.
As a result, HDFG output is sensitive to the relative overlap of potential energy surfaces.
If the potential energy surface is not significantly modulated along the vibrational coordinate of interest, then output is dominated by $A$ term resonances, or Franck-Condon contributions.
However, $B$ term resonances, or Herzberg-Teller contributions, dominate in centrosymmetric media, as $A$ term resonances vanish in centrosymmetric media through parity constraints.
A treatment of inhomogeneous broadening shows that for positively correlated vibrational and vibronic resonances, HDFG linenarrows the spectral output, eliminating inhomogeneous broadening. 
As a result, HDFG possesses site-selective properties analogous to those of coherent Raman spectroscopies, which can allow for a thorough analysis of vibronic coupling schemes in isotropic systems.
We show that HDFG provides alternative methods to extract the hyper-Raman hyperpolarizability of infrared active vibrations. 
Since HDFG is approachable as a two-beam experiment, analogous to vibrational sum-frequency generation, and phase-matchable in media with normal dispersion, HDFG should be applicable towards understanding vibrational spectroscopy and dynamics, e.g. single quantum coherence free induction decay dynamics, in a variety of relevant molecular and materials systems.

@@

\section{Supplemental Information}
See the supporting information for details on deriving closed-form solutions to Herzberg-Teller integrals, dependence of the HDFG spectrum on $\Delta$, and integration of \autoref{contour}.

>>>>

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{taxonomy.png}
	\caption{
		Some spectroscopic methods for investigating (top) electronic-vibrational coupling and (bottom) e-v-x (x: e, v, e-v, e-v-e) coupling through different order nonlinear processes.
		The light-matter interactions are depicted using Wave Mixing Energy Level (WMEL) diagrams.\cite{RN286}
		Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. 
		The red box outlines the e-v coupling techniques (HDFG, HSFG) relevant to this manuscript.
	} 
	\label{fig:comparisonwmel}
\end{figure*}

@@

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{drsive_spectrum.png}
	\caption{Contributions of $A, B$ (\autoref{ABterms_DR}) to HDFG spectrum for a simple two harmonic well system.
		(a) Potential energy surfaces for a two-well system, (b) 1D HDFG spectrum for $\Delta = 0.5$, $\omega_1 = \omega_{g1, g0}$, (c, d) real and imaginary parts of terms plotted in (b), respectively.
		The spectra are referenced to $\max{(A+B)}$. 
		It is assumed that $\omega_2 = \omega_3$.
		The vibrational states on the $|g)$ and $|e)$ manifolds are spaced 1600 cm$^{-1}$ apart, with the vibronic states $\ket{e,\tilde{v}'}$ given linewidths of 300 cm$^{-1}$ and $\omega_{eg}$ = 30000 cm$^{-1}$.
		Dotted lines in (b) denote the $\tilde{v}'$ = 0, 1, 2 vibronic resonances. 
		The vibronic one and two photon absorption operators are scaled such that $\abs{B}/\abs{A} \sim$ 0.1, following Chung and Ziegler. \cite{Ziegler1988}
		The Franck-Condon and Herzberg-Teller contributions to each of these terms are found in \autoref{T:contributions}.}
	\label{fig:doubres_spec}
\end{figure*}

@@

Since $\Delta$ can take on a variety of values given a specific system, in the supporting information we demonstrate the impact of $\Delta$ on the HDFG response using previously reported values for some Raman active vibrations in \textit{trans}-hexatriene and UiO-68 metal organic frameworks (Figure S2). \cite{Myers1982, Brennan2024}
Clearly, specific $\Delta$ and $\omega_{g1, g0}$ values have a significant impact on the HDFG spectral profile.
Indeed, for some values of $\Delta$, it appears that $B$ term resonances dominate over $A$ term resonances, which can allow for significant non-Condon behavior in pre-resonant HDFG response.

In the post-resonance region, the amplitude takes on an unexpected lineshape.
This arises due to interference between the real and imaginary parts of $\gamma$ (\autoref{fig:doubres_spec}c, d).
The abnormal lineshapes in both terms arises due to the sign of the Franck-Condon and Herzberg-Teller integrals that drive the output of the vibronic transitions for $\Delta = 0.5$ (\autoref{fcht}), in turn driving significant interference in the overall $\abs{A+B}$ output.
Such effects have been seen elsewhere in spontaneous, surface enhanced hyper-Raman studies. \cite{Milojevich2011}

@@

\section{Conclusions}\label{conclusion}
Coherent vibrational, hyper-Raman coherent four wave mixing spectroscopies are identified and discussed.
Hyper difference frequency generation (HDFG) possesses infrared absorption and vibronic hyper-Raman scattering selection rules.
As a result, HDFG output is sensitive to the relative overlap of potential energy surfaces.
If the potential energy surface is not significantly modulated along the vibrational coordinate of interest, then output is dominated by $A$ term resonances, or Franck-Condon contributions.
However, $B$ term resonances, or Herzberg-Teller contributions, dominate in centrosymmetric media, as $A$ term resonances vanish in centrosymmetric media through parity constraints.
A treatment of inhomogeneous broadening shows that for positively correlated vibrational and vibronic resonances, HDFG linenarrows the spectral output, eliminating inhomogeneous broadening. 
As a result, HDFG possesses site-selective properties analogous to those of coherent Raman spectroscopies, which can allow for a thorough analysis of vibronic coupling schemes in isotropic systems.
We show that HDFG provides alternative methods to extract the hyper-Raman hyperpolarizability of infrared active vibrations. 
Since HDFG is approachable as a two-beam experiment, analogous to vibrational sum-frequency generation, and phase-matchable in media with normal dispersion, HDFG should be applicable towards understanding vibrational spectroscopy and dynamics, e.g. single quantum coherence free induction decay dynamics, in a variety of relevant molecular and materials systems.

@@

\section{Supplemental Information}
See the supporting information for details on deriving closed-form solutions to Herzberg-Teller integrals, HDFG spectrum dependence on $\Delta$, and evaluation of \autoref{contour}.

### 2024-09-18 si.tex
edits

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals, in the limit of no Duschinsky coupling, when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933, Duschinsky1937}
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $H_e$ and $H_g$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
Franck-Condon factors and Herzberg-Teller integrals are often evaluated in position space using known solutions to the Harmonic Oscillator Hamiltonian, 
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N(m) H_m(q) \exp(-q^2/2),
\end{equation}
where $N(m) = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the m$^{th}$ Hermite Polynomial.\cite{RN230}
We begin by generating an expression for $\mel{\tilde{n}}{q^k}{m}$, as this will resolve into Franck-Condon factors for $k = 0$ and Herzberg-Teller integrals for $k=1$.
As in the main text, $\{\ket{\tilde{a}}\}$ are vibrational states on the first excited state $|e)$ and $\{\ket{b}\}$ are vibrational states on the ground state $|g)$.

@@

\section{Impact of $\Delta$ on $A,B$ Terms in HDFG Spectra}
In the main text, the potential well offset ($\Delta$) was taken as $\Delta = 0.5$.
Here, we vary $\Delta$ to investigate its impact on the structure of the $\abs{A+B}$ spectrum.
The relevant Franck-Condon factors and Herzberg-Teller integrals, assuming identical curvature and no Duschinsky coupling, are plotted in \autoref{fig:fcht} as a function of offset, $\Delta$.

It is clear from \autoref{fig:fcht} that the HDFG spectrum will have a significant dependence upon $\Delta$, even for small deviations from complete overlap. 
To this end, we first investigate the impact of $\Delta$ on a vibrational mode with a fundamental frequency 1600 cm$^{-1}$ and linewidth of 400 cm$^{-1}$ (\autoref{fig:chgdelta}).

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/changedelta.png}
	\caption{Plots of select (a) Franck-Condon factors and (b) Herzberg-Teller integrals relevant to this manuscript as a function of offset, $\Delta$.} 
	\label{fig:chgdelta}
\end{figure}
\section{Evaluating impact of inhomogeneity on HDFG output}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated using Contour integration.

>>>>

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals, in the limit of no Duschinsky coupling, when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933, Duschinsky1937}
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $H_e$ and $H_g$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
Franck-Condon factors and Herzberg-Teller integrals are often evaluated in position space using known solutions to the Harmonic Oscillator Hamiltonian, 
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N(m) H_m(q) \exp(-q^2/2),
\end{equation}
where $N(m) = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the m$^{th}$ Hermite Polynomial.\cite{RN230}
We begin by generating an expression for $\mel{\tilde{n}}{q^k}{m}$, as this will resolve into Franck-Condon factors for $k = 0$ and Herzberg-Teller integrals for $k=1$.
The relevant Franck-Condon factors and Herzberg-Teller integrals, assuming identical curvature and no Duschinsky coupling, are plotted in \autoref{fig:fcht} as a function of offset, $\Delta$.
As in the main text, $\{\ket{\tilde{a}}\}$ are vibrational states on the first excited state $|e)$ and $\{\ket{b}\}$ are vibrational states on the ground state $|g)$.

@@

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/changedelta.png}
	\caption{The impact of offset, vibrational resonance, and linewidth on HDFG spectra. Parameters for (a,b) follow from Myers et al., whereas parameters for (c,d) follow from Brennan et al.\cite{Myers1982, Brennan2024}} 
	\label{fig:chgdelta}
\end{figure}

\section{Impact of $\Delta$ on $A,B$ Terms in HDFG Spectra}
In the main text, the potential well offset ($\Delta$) was taken as $\Delta = 0.5$.
Here, we vary $\Delta$ to investigate its impact on the structure of the $\abs{A+B}$ spectrum.

It is clear from \autoref{fig:fcht} that the HDFG spectrum will have a significant dependence upon $\Delta$, even for small deviations from complete overlap. 
To this end, we plot the impact of resonance frequency and offset on the HDFG spectrum \autoref{fig:chgdelta}, using values reported by Myers et al. and Brennan et al. \cite{Myers1982, Brennan2024}
In this section, we take $\omega_{eg}$ = $37340$ cm$^{-1}$ and vary the offset ($\Delta$), vibrational frequency ($\omega_{g1,g0}$), and linewidth ($\Gamma$) as noted in the figure.

\section{Evaluating impact of inhomogeneity on HDFG output}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated using Contour integration.
